---
title: "DS607-Project1"
author: "George Cruz"
date: "9/17/2020"
output: html_document
---
## Project Outline
In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:

-Player’s Name 
-Player’s State, 
-Total Number of Points, 
-Player’s Pre-Rating, 
-Average Pre Chess Rating of Opponent

### Setting up supporting functions
To deal with the data's ETL process, I will create several r functions to avoid code repetition.  These functions will be listed here.
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r echo=TRUE}
#returns a list of rows in a data frame
rows = function(tab) lapply(
  seq_len(nrow(tab)),
  function(i) unclass(tab[i,,drop=F])
)

# creates an empty data frame
create_empty_table <- function(num_rows, num_cols, type_vec) {
  frame <- data.frame(matrix(NA, nrow = num_rows, ncol = num_cols))
  for(i in 1:ncol(frame)) {
    print(type_vec[i])
    if(type_vec[i] == 'numeric') {frame[,i] <- as.numeric(frame[,i])}
    if(type_vec[i] == 'character') {frame[,i] <- as.character(frame[,i])}
    if(type_vec[i] == 'logical') {frame[,i] <- as.logical(frame[,i])}
    if(type_vec[i] == 'factor') {frame[,i] <- as.factor(frame[,i])}
  }
  return(frame)
}

# removes the pipe character and trims the strings
clean_text <- function(text ) { 
  text <- str_trim(gsub("\\|", "", text))
  return(text)
}

#gets the rating of an opponent based on the result
get_opp_rating <- function( df, result) { 
  #extract player's ID from result 
  opp_num <- as.numeric(str_extract_all(result, "\\d+"))
  opp_rating <- df[opp_num,]$pre_rating
  
  return(opp_rating)
}

# gets the average opponent rating by looking at each 
# opponent's ratings
get_avg_opponent_rtg <- function( df, id) { 
    opp_ratings <- c( get_opp_rating(df, df[id, ]$round1), 
                      get_opp_rating(df, df[id, ]$round2),
                      get_opp_rating(df, df[id, ]$round3),
                      get_opp_rating(df, df[id, ]$round4),
                      get_opp_rating(df, df[id, ]$round5),
                      get_opp_rating(df, df[id, ]$round6),
                      get_opp_rating(df, df[id, ]$round7))
    return(mean(opp_ratings, na.rm = TRUE))
}
```

## Extract the data
The data lives in my github, and follows this format: 
`-----------------------------------------------------------------------------------------`
` Pair | Player Name                     |Total|Round|Round|Round|Round|Round|Round|Round| `
` Num  | USCF ID / Rtg (Pre->Post)       | Pts |  1  |  2  |  3  |  4  |  5  |  6  |  7  | `
`-----------------------------------------------------------------------------------------`
`    1 | GARY HUA                        |6.0  |W  39|W  21|W  18|W  14|W   7|D  12|D   4|`
`   ON | 15445895 / R: 1794   ->1817     |N:2  |W    |B    |W    |B    |W    |B    |W    |`
`-----------------------------------------------------------------------------------------`

### Getting the data
We will first get the data into a dataframe called tournament.
We will remove the lines with only dashes

```{r}
#getting the data from the file
theURL <- "https://raw.githubusercontent.com/georg4re/DS607/master/project1/tournamentinfo.txt"

#using read.fwf will help me with the fixed width nature of the document
tournament <- read.fwf(theURL, width = c(7,34,6,6,6,6,6,6,6,6))

#removing the separating lines between player information
nEnd <- nrow(tournament)
if ((nEnd -1) %% 3 > 0 ) { 
  nEnd = nEnd - ((nEnd -1) %% 3) + 1
} 

removeRows = seq(from=1, to= nEnd, by=3)
tournament <- tournament %>% slice(-removeRows)

# Remove the title values
clean_tournament <- tournament %>% slice(-c(1,2))
head(clean_tournament,10)
```

## Transform the data into a player related dataframe

With the pieces in place, we can now transform the data into a player-centric dataframe.

### Create the empty receiving table: player_data
```{r}

#set the column types
types <- c("numeric","character","character","character","character","character","character","character","character","character", "character", "character")

player_data <- create_empty_table(0, 12, types)
```

Let's iterate thru the original data to get our values into player_data
```{r}
line1 <- vector()
line2 <- vector()

for (i in 1 : nrow(clean_tournament)) {
  # If it's the second line, add the player info to the table
  if ((i %% 2) == 0) {
    line2 <- clean_tournament[i, ]

    player_data <- rbind(player_data, c(as.numeric(clean_text(line1$V1)), clean_text(line1$V2),
                                        clean_text(line1$V3), clean_text(line1$V4),
                                        clean_text(line1$V5), clean_text(line1$V6),
                                        clean_text(line1$V7), clean_text(line1$V8),
                                        clean_text(line1$V9), clean_text(line1$V10),
                                        clean_text(line2$V1), clean_text(line2$V2)))
  } 
  else { 
    line1 <- clean_tournament[i, ]
  }
}


titles <- c("player_num","name","total_pts","round1","round2","round3","round4","round5","round6","round7", "state", "ID_Rtg")

colnames(player_data) <- titles
#Ensure player_num is numeric 
player_data$player_num <- as.numeric(as.character(player_data$player_num))
head(player_data,5)
```

We now have the data somewhat in place.  We can start cleaning it up by extracting the values we need.

### Extract the ID, Pre Rating and Post Rating into their own columns 
```{r}
extracted_id_rtg <- data.frame( player_num= player_data$player_num,
                                uscf_id = str_extract(player_data$ID_Rtg, "\\d+"),
                                pre_rating = as.numeric(str_extract(player_data$ID_Rtg, 
                                                                    "(?<=R: ).\\d+(?-)")),
                                post_rating = as.numeric(
                                  str_extract(
                                    str_replace_all(player_data$ID_Rtg,"->","P:"),
                                    "(?<=P:).\\d+")),
                                stringsAsFactors=FALSE)

player_data <- merge(x= player_data, y = extracted_id_rtg, by="player_num", all.x = TRUE)
head(player_data,5)


```
### Calculate Opponents Pre rating
Now, we will calculate the player's opponents' average pre-ratings and merge it to our 
player data table
```{r}

avg_opp_rtg <- data.frame( player_num= "1",
                       avgOppRating= as.numeric(get_avg_opponent_rtg(player_data,1)),
                       stringsAsFactors=FALSE)

for (i in 2:nrow(player_data)) { 
  avg_opp_rtg <- rbind( avg_opp_rtg, c(as.character(i), 
                        as.numeric(get_avg_opponent_rtg(player_data,i))
                        ))  
}
player_data <- merge(x= player_data, y = avg_opp_rtg, by="player_num", all.x = TRUE)
head(player_data,5)

```


### Producing the resulting data frame

```{r}
final_df <- data.frame(Name= player_data$name, 
                       State= player_data$state, 
                       total_pts= player_data$total_pts, 
                       pre_rating= player_data$pre_rating, 
                       opp_rating= as.numeric(as.character(player_data$avgOppRating)))
head(final_df,5)
```

### Produce Final CSV file
```{r}
write.csv(final_df, file = "ds607-gc-project1.csv")
```

9/19/2020


