---
title: "DS607-HW7"
author: "George Cruz"
date: "10/10/2020"
output: html_document
---

## Data 607 Homework 7

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.
Take the information that you’ve selected about these three books, and separately create three files which
store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”,
“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you
create each of these files “by hand” unless you’re already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(httr)
library(rjson)
library(RCurl)
library(arsenal)  #for comparison purposes
```

### HTML Data
```{r}
html_source<-readHTMLTable(getURL("https://raw.githubusercontent.com/georg4re/DS607/master/data/books.html"))
html_source<-lapply(html_source[[1]], function(x) {unlist(x)})
html_data<-as.data.frame(html_source)
head(html_data)
```

### XML Data
```{r}
xml_source<-xmlInternalTreeParse(getURL("https://raw.githubusercontent.com/georg4re/DS607/master/data/books.xml"))
xml_source<-xmlSApply(xmlRoot(xml_source), function(x) xmlSApply(x, xmlValue))
#Need to flip the rows and columns
xml_source <- t(xml_source)
xml_data<-data.frame(xml_source, row.names = NULL)
head(xml_data)
```

### JSON Data
```{r}
json_source <- fromJSON(file = "https://raw.githubusercontent.com/georg4re/DS607/master/data/books.json")
json_source <- lapply(json_source, function(x) {
  unlist(x)
})
json_data <- as.data.frame((do.call("rbind", json_source)))
head(json_data)
```

### Comparing the Data
```{r}
comparedf(json_data, xml_data)

```

When comparing the dataframes, we see the variable names and values are the same between JSON and XML,  Due to the way we named them 

HTML variable names are different because we used case titles for the columns 
```{r}
summary(comparedf(html_data, json_data))
```

