---
title: "DS607-HW8"
author: "George Cruz"
date: "`r Sys.Date()`"
output: openintro::lab_report
---
# Assignment – Web APIs

The task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it into an R DataFrame.

### Setting up

I will store the api-key value in an environment variable by using Sys.setenv

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(httr)
library(jsonlite)

#Storing the API Key in an environment file
api_key <- Sys.getenv("api_key")
```

### Connecting to the API
To connect to the API, we append the api-key as a query string value to the API URL.

```{r}
#To connect to the API, we compose the url
url <- paste("https://api.nytimes.com/svc/movies/v2/reviews/picks.json?api-key=",
             api_key, sep="")
res <- GET(url)
data <- fromJSON(rawToChar(res$content))

my_data_frame <- data$results
```


### Analyzing the Data
We begin by taking a look at the data provided by the API:

```{r}

summary(my_data_frame)
```


```{r}
knitr::kable(head(my_data_frame))
```

Because we used the Movies API to only get critic's picks, we see the pick column value is always one. 

Let's count the movies by rating: 
```{r}
movies_ratings <- my_data_frame %>%
  count(mpaa_rating) %>%
  mutate(p = n /sum(n)) 

movies_ratings$mpaa_rating[1] <- "No Rating"
knitr::kable(movies_ratings)
```


```{r code-chunk-label}
ggplot(data=my_data_frame, aes(x=mpaa_rating, stat="count")) + 
 geom_bar()
```

### Conclusion
It appears that most of the NY Times Critic's picks are not rated by the MPAA.  Further analysis could be done to try to infer the probability of a movie being picked by a critic based on its description (to try to determine its genre and other descriptive features).



...

