---
title: "DS607 - Homework 3"
author: "George Cruz"
date: "`r Sys.Date()`"
output: openintro::lab_report
---
# Homework 3

### 1. Using the 173 majors listed 
In fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
```{r load-packages, message=FALSE}
library(tidyverse)
```


```{r}
theURL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
all_majors <- read.csv(file=theURL, fileEncoding="UTF-8-BOM")
data_or_stats <- subset(all_majors, grepl("DATA|STATISTICS", Major))
data_or_stats
```

### 2. Write code that transforms the data below:

```{r code-chunk-label}

vector_1 <- c("bell pepper", "bilberry", "blackberry", "blood orange")
vector_2 <- c("blueberry", "cantaloupe", "chili pepper", "cloudberry")
vector_3 <- c("elderberry", "lime", "lychee", "mulberry")
vector_4 <- c("olive", "salal berry")
main_vector <- c(vector_1, vector_2, vector_3, vector_4)
main_vector
```

### 3. Describe, in words, what these expressions will match:
```{r, results='hide'}
lista <- c("abba", "hello\2\1","aaaabbbcccdddde", "aabbmmmmmjmmkmkkk","banana", "amanaplanacanalpanama","civic", "racecar")
str_view(lista, "(.)\\1\\1")
str_view(lista, "(.)(.)\\2\\1")
str_view(lista, "(..)\\1")
str_view(lista, "(.).\\1.\\1")
str_view(lista, "(.)(.)(.).*\\3\\2\\1")

```

1. "(.)\1\1"  Match any three repeated characters
2. "(.)(.)\\2\\1 Match any symmetrical 4 characters where 1st and 4th characters are the same and 2,3 are the same.
3. (..)\1 Match any repeated characters separated by 1 character.
4. "(.).\\1.\\1" match any three repeated characters separated by the same character
5. "(.)(.)(.).*\\3\\2\\1" match palyndromes of 6 of more letters

### 4.Construct regular expressions to match words that:

Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r, results='hide'}
words <- c("civic", "church", "eleven")
str_view(words, "^(.).*\\1$")
str_view(words, "(.)(.).*\\1\\2")
str_view(words, "(.).\\1.\\1")

```


